### 2020-06-26 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/paper/2-monotone-quintic.tex
Changed some directory names and fixed some typos in code files. Still working on the paper.

\item{} {The necessity of Steps 2 and 4 follows directly from
  the fact that $f$ is $C^2$. The following Steps 7--13 coincide with
  a simplified condition for quintic monotonicity that reduces to one
  of cubic positivity studied by Schmidt and He{\ss} [1988]. Given
  $\alpha$, $\beta$, $\gamma$, and $\delta$ as defined by Schmidt and
  He{\ss}, monotonicity results when $\alpha \geq 0$, $\delta \geq 0$,
  $\beta \geq \alpha - 2 \sqrt{\alpha \delta}$, and $\gamma \geq
  \delta - 2 \sqrt{\alpha \delta}$.  Step 4 checked for $\delta < 0$,
  Step 8 checks $\alpha < 0$, Step 10 checks $\beta < \alpha - 2
  \sqrt{\alpha \delta}$, and Step 11 checks $\gamma < \delta - 2
  \sqrt{\alpha \delta}$. If none of the monotonicity conditions were
  violated, then the order six piece is monotone and Step 12
  concludes.}

@@

Given the minimum magnitude curvature nature of the initial derivative
estimates, it is desirable to make the smallest necessary changes to
the initial interpolating spline $Q$ while enforcing monotonicity. In
practice a quasi-bisection search is used in place of solely shrinking
$DQ$ and $D^2Q$, adding an additional step that increases previously
shrunken derivative values by a smaller increment than the previous
decrement. The goal is to converge on the boundary of the monotone region
in the $(\tau_1$, $\alpha$, $\beta$, $\gamma)$ space. As long as a step size
schedule is set that allows the value zero to be obtained in a fixed
number of computations, this has no effect on computational
complexity. Notably, since {\tt IS\_MONOTONE} handles both nondecreasing and
nonincreasing simultaneously by taking into account the sign of $v$,
Algorithm 3 produces $Q(x)$ that is monotone increasing (decreasing) over
exactly the same intervals that the data $(X_i$, $Y_i)$ is monotone
increasing (decreasing).

>>>>

\item{} {The necessity of Steps 2 and 4 follows directly from
  the fact that $f$ is $C^2$. The following Steps 7--13 coincide with
  a simplified condition for quintic monotonicity that reduces to one
  of cubic positivity studied by Schmidt and He{\ss} [1988]. Given
  $\alpha$, $\beta$, $\gamma$, and $\delta$ as defined by Schmidt and
  He{\ss}, monotonicity results when $\alpha \geq 0$, $\delta \geq 0$,
  $\beta \geq \alpha - 2 \sqrt{\alpha \delta}$, and $\gamma \geq
  \delta - 2 \sqrt{\alpha \delta}$.  Step 4 checked for $\delta < 0$,
  Step 8 checks $\alpha < 0$, Step 10 checks $\beta < \alpha - 2
  \sqrt{\alpha \delta}$, and Step 11 checks $\gamma < \delta - 2
  \sqrt{\alpha \delta}$. If none of the monotonicity conditions are
  violated, then the order six piece is monotone and Step 12
  concludes.}

@@

Given the minimum magnitude curvature nature of the initial derivative
estimates, it is desirable to make the smallest necessary changes to
the initial interpolating spline $Q$ while enforcing monotonicity. In
practice a quasi-bisection search is used in place of solely shrinking
$DQ$ and $D^2Q$. This algorithm is outlined below. The goal is to
converge on the boundary of the monotone region in the $(\tau_1$,
$\alpha$, $\beta$, $\gamma)$ space.

%% ===================================================================
%%              Algorithm 4: Quasi-bisection search
\vskip 5mm
{\parindent 0mm
{\bf Algorithm 4:} {\tt QUASI-BISECTION}$\bigl(x_0, x_1, f \bigr)$

where $x_0$, $x_1 \in {\bbb R}$, $x_0 < x_1$, and $f$ is an order six
polynomial that is nonmonotone on the interval $[x_0$, $x_1]$. Defines
a stepping procedure for function derivative values that ensures $f$
can be made monotone in a fixed number of steps given a relative
precision $e \in {\bbb R}$.

}
{\parindent=6mm
\item{} Set a starting {\tt step\_size} and save the initial value of $f$.
\item{} $\hat f := f$
\item{} {\tt step\_size} $:= 1/2$
\item{} {\tt shrinking} $:=$ {\tt TRUE}
\item{} {\tt do while } $\big( ($ {\tt step\_size} $ > e)$ {\tt or not
  IS\_MONOTONE}$(x_1$, $x_2$, $f)$ $\big)$
\item{} \codent {\tt if shrinking then step\_size} $:=$ {\tt step\_size} $/ 2$

}
\vskip 5mm
%% ----------------------------------------------------------------------

As long as a step size schedule is
set that allows the value zero to be obtained in a fixed number of
computations, this has no effect on computational complexity. Notably,
since {\tt IS\_MONOTONE} handles both nondecreasing and nonincreasing
simultaneously by taking into account the sign of $v$, Algorithm 3
produces $Q(x)$ that is monotone increasing (decreasing) over exactly
the same intervals that the data $(X_i$, $Y_i)$ is monotone increasing
(decreasing).

### 2020-06-26 Research_[2020-05-18]_Piecewise_Monotone_Quintic_Fortran/paper/4-analysis.tex
Changed some directory names and fixed some typos in code files. Still working on the paper.

>>>>

\topinsert
\centerline{\epsfxsize=4truein \epsffile{vis/1-sensitivity.eps}}
{\narrower\noindent\rmVIII Fig.\ 1.
A demonstration of the quadratic facet model's sensitivity to small
data perturbations. In this example two quadratic functions $f_1(x) =
x^2$ over $\{1$,$2$,$5/2\}$ and $f_2(x) = (x-2)^2 + 6$ over
$\{5/2$,$3$,$4\}$ have the same curvature and equal value at
$x=5/2$. Given the five points above Algorithm 2 produces the slope of
the solid blue line because $f_1$ and $f_2$ have equal curvature.
However, perturbing the curvature of $f_2$ down by the machine
precision at values $x=\{3$, $4\}$ causes Algorithm 2 to produce the
slope of the red dashed line at $x = 5/2$. This is the nature of a
facet model and a side effect of associating data with local facets.
\par}
\endinsert
